We are not the 99 percent: the Local Group as an outlier in galaxy pairs' satellite distributions

### 2017-10-27 paper/paper.tex
caption

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[No Title]{No Title}
\author[J.E. Forero-Romero \& V. Arias]
{Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co},
Ver\'onica Arias$^1$\\
%%
$^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
}

@@

\begin{abstract}
We quantify the joint spatial distribution of the brightest satellites
around the Milky Way (MW) and M31, the dominant galaxies in the Local
Group (LG).
We use the Illustris-1 and ELVIS simulations to show that the
MW satellite distribution is significantly thinner and more oblate
than the expectations from the Lambda Cold Dark Matter (LCDM)
paradigm. In contrast, the satellites in M31 are fully consistent with the
average expectations from simulations.
We also find that the satellite planes in the MW and M31 are almost perpendicular to the line
connecting the two galaxies. While none of the pairs in the simulations show
a similar degree of alignment, these alignments are consistent with a
random uniform distribution, making the observations
consistent with theoretical expectations.
The special satellite distribution in the Milky Way makes the Local
Group an atypical pair in the cosmological context provided by
LCDM simulations. 
We estimate that only $0.5\%$ to $1\%$ of the pairs with isolation
and dynamical characteristics similar to the LG are expected to have
satellite distributions with the same degree of atipicality, making
the LG a $3\sigma$ outlier in a sample of simulated galaxy pairs.
\end{abstract}

@@

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{gaussian_model_illustris_M31.pdf}
\includegraphics[width=0.45\textwidth]{gaussian_model_illustris_MW.pdf}
\caption{Correlations between the normalized values for the plane
  width, $c/a$ ratio and $b/a$ ratio. 
The star corresponds to the LG, black circles with errorbars are the
results from simulations and the gray cloud is the resul from the
multivariate gaussian model.
Upper/lower row summarizes the results for M31/M31.
This simplified description allows us to quantify how atypical is the
LG compared to the simulation results. 
\label{fig:correlations}}
\end{figure*}

@@

\subsection{s}

\subsection{Satellite distributions in the Illustris Simulation}

In the previous section we saw how the MW cannot be concealed with its
randomized satellite distribution, while M31 is fully consistent with
it. 
We now quantify whether the MW and M31 are consistent with the
expectations from the Illustris Simulation.
We show here the results from the \rank\ sample. 
The results from the \boot\ sample ara avaialble in the appendix.

@@

\begin{table*}
  \centering
\begin{tabular}{lll}
\hline\hline
Symbol & Units & Description\\\hline
$\hat{r}_{AB}$& & Unit vector along the direction connecting two
dominant galaxies\\
$N_s$ & & Number of satellites\\
$a > b> c$ & kpc & Inertia tensor eigenvalues. \\
$\hat{I}_1$, $\hat{I}_2$, $\hat{I}_3$ & & Inertia tensor eigenvectors. \\
$\sigma_s$ & kpc & Ellipsoid width\\
\hline\hline
\end{tabular}
  \caption{Overview of the parameters computed for each central galaxy
    and its satellite system.
  \label{tab:models}}
\end{table*}

@@

%MW mass
%https://arxiv.org/abs/1407.1078
\appendix

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_ranked_elvis_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_elvis_width.pdf}\\
\includegraphics[width=0.30\textwidth]{scatter_ranked_elvis_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_elvis_ca_ratio.pdf}\\
\includegraphics[width=0.30\textwidth]{scatter_ranked_elvis_ba_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_elvis_ba_ratio.pdf}\\
\includegraphics[width=0.30\textwidth]{scatter_ranked_elvis_mu.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_elvis_mu.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $b/a$ axis ratio. In this case both the MW and M31
are consistent with the results of a spherical distribution and the
simulations. 
\label{fig:scatter_ba_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{gaussian_model_elvis_M31.pdf}
\includegraphics[width=0.45\textwidth]{gaussian_model_elvis_MW.pdf}
\caption{Correlations between the normalized values for the plane
  width, $c/a$ ratio and $b/a$ ratio. 
The star corresponds to the LG, black circles with errorbars are the
results from simulations and the gray cloud is the resul from the
multivariate gaussian model.
Upper/lower row summarizes the results for M31/M31.
This simplified description allows us to quantify how atypical is the
LG compared to the simulation results. 
\label{fig:correlations}}
\end{figure*}

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[LG as an outlier]{We are not the 99 percent: the Local Group as an
  outlier in galaxy pairs' satellite distributions}
\author[J.E. Forero-Romero \& V. Arias]
{Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co},
Ver\'onica Arias$^1$\\
%%
$^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
}

@@

\begin{abstract}
We quantify the joint spatial distribution of the brightest satellites
around the Milky Way (MW) and M31, the dominant galaxies in the Local
Group (LG).
We use the Illustris-1 and ELVIS simulations to show that the
MW satellite distribution is significantly thinner and more oblate
than the expectations from the Lambda Cold Dark Matter (LCDM)
paradigm. In contrast, the satellites in M31 are fully consistent with the
average expectations from simulations.
We also find that the satellite planes in the MW and M31 are almost perpendicular to the line
connecting the two galaxies. While none of the pairs in the simulations show
a similar degree of alignment, these alignments are consistent with a
random uniform distribution, making the observations
consistent with theoretical expectations.
The special MW satellite distribution makes the LG an atypical pair in
the cosmological context provided by LCDM simulations. 
We estimate that only $0.5\%$ to $1\%$ of the pairs with isolation
and dynamical characteristics similar to the LG are expected to have
satellite distributions with the same degree of atipicality, making
the LG a $3\sigma$ outlier in a sample of simulated galaxy pairs.
\end{abstract}

@@

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{gaussian_model_illustris_M31.pdf}
\includegraphics[width=0.45\textwidth]{gaussian_model_illustris_MW.pdf}
\caption{Correlations between the multivariate gaussian model built on
  the normalized values for the plane width $w$, $c/a$ ratio and $b/a$ ratio. 
Left/right panel correspond to M31/MW. 
The contour levels in the 2D histograms correspond to the $1\sigma$,
$2\sigma$ and $3\sigma$ contours in two dimensions. 
The dahsed vertical lines in the histograms along the diagonal
correspond to the $1\sigma$ boundaries in one dimension.
The results for the gaussian model are built from $10^6$ point
realizations in the six-dimensional space spaned by the variables of
interest. 
The correlation matrix and the mean values are computed from the
results in the Illustris-1 simulation.
The cross indicates the LG values.
This plot clearly shows how the M31 results are well within the
expectations from simulations while MW has an unusual low value for
the plane width and the $c/a$ axis ration.
Similar results from the ELVIS data are presented in Figure
\ref{fig:correlations_elvis}.
\label{fig:correlations_illustris}}
\end{figure*}

@@

\subsection{Plane alignment}

@@

@@

%MW mass
%https://arxiv.org/abs/1407.1078
\newpage
\appendix

\section{Results from ELVIS}
\begin{figure*}
\centering
\includegraphics[width=0.24\textwidth]{scatter_ranked_elvis_width.pdf}
\includegraphics[width=0.24\textwidth]{scatter_ranked_elvis_ca_ratio.pdf}
\includegraphics[width=0.24\textwidth]{scatter_ranked_elvis_ba_ratio.pdf}
\includegraphics[width=0.24\textwidth]{scatter_ranked_elvis_mu.pdf}
\includegraphics[width=0.24\textwidth]{scatter_norm_ranked_elvis_width.pdf}
\includegraphics[width=0.24\textwidth]{scatter_norm_ranked_elvis_ca_ratio.pdf}
\includegraphics[width=0.24\textwidth]{scatter_norm_ranked_elvis_ba_ratio.pdf}
\includegraphics[width=0.24\textwidth]{scatter_norm_ranked_elvis_mu.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $b/a$ axis ratio. In this case both the MW and M31
are consistent with the results of a spherical distribution and the
simulations. 
\label{fig:scatter_ba_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{gaussian_model_elvis_M31.pdf}
\includegraphics[width=0.45\textwidth]{gaussian_model_elvis_MW.pdf}
\caption{Correlations between the normalized values for the plane
  width, $c/a$ ratio and $b/a$ ratio. 
The star corresponds to the LG, black circles with errorbars are the
results from simulations and the gray cloud is the resul from the
multivariate gaussian model.
Upper/lower row summarizes the results for M31/M31.
This simplified description allows us to quantify how atypical is the
LG compared to the simulation results. 
\label{fig:correlations_elvis}}
\end{figure*}
